---
title: "CSDE 502 HW 7"
author: "Colin Baynes (cbaynes@uw.edu)"
date: "2/22/2021"
output:
  bookdown::html_document2:
    number_sections: yes
    self_contained: yes
    code_folding: hide
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: no
  html_document:
    toc: yes
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, include=F}
rm(list=ls())
```

# Problem 1
```{r, include=F}
#Load packages
library(tidyverse)
library(magrittr)
library(knitr)
library(kableExtra)
library(readstata13)
library(haven)
library(curl)
library(ggplot2)

```
*Note: we obtained the dataset from http://staff.washington.edu/phurvitz/csde502_winter_2021/data/21600-0001-Data.dta.zip. To reproduce this code without resource from local computer the user should obtain the full public Add Health Data Set from : http://www.cpc.unc.edu/projects/addhealth.*

I am using variables from the publicly available data set Add Health to complete this assignment. The variables I am using from Add Health concerns participants' number of times per week doing chores around the house, doing hobbies, watching television or playing video games, playing sports and doing exercise. To look at these variables we generate a new data frame with this information from participants. We made sure to include appropriate labels for the new data frame and its variables.
```{r, warning=F, include=F}
setwd('C:/Users/Cbayn/Dropbox/CSDE 502/week 7')
ah_wave1_v1<- read.dta13("21600-0001-Data.dta")

#creating data frame include five variables on participants' times per week of the following activities: (1) chores, (2) hobbies, (3) television or video games, (4) sports, and (5) exercise
ah_wave1_v1_subset <- subset(ah_wave1_v1, select = c(AID, H1DA1, H1DA2, H1DA3, H1DA5, H1DA6))

#creating brief label attribute for the data frame that says what the subsetted data frame is:
attributes(ah_wave1_v1_subset)$label <- "Add Health Wave I Cohort Members Weekly Activities"
ah_wave1_v1_subset %>% attributes() %>% extract("label")
#create as.factor variables
ah_wave1_v1_subset$UID<-as.factor(ah_wave1_v1_subset$AID)
ah_wave1_v1_subset$Chores<-as.factor(ah_wave1_v1_subset$H1DA1)
ah_wave1_v1_subset$Hobbies<-as.factor(ah_wave1_v1_subset$H1DA2)
ah_wave1_v1_subset$TV<-as.factor(ah_wave1_v1_subset$H1DA3)
ah_wave1_v1_subset$Sports<-as.factor(ah_wave1_v1_subset$H1DA5)
ah_wave1_v1_subset$Exercise<-as.factor(ah_wave1_v1_subset$H1DA6)
ah_wave1_v1_subset <- subset(ah_wave1_v1_subset, select = c(UID, Chores, Hobbies, TV, Sports, Exercise))
#attributes for the variables
attributes(ah_wave1_v1_subset$UID)$label <- "Participant Unique Identifier"
attributes(ah_wave1_v1_subset$Chores)$label <- "Occasions per week of doing chores"
attributes(ah_wave1_v1_subset$Hobbies)$label <- "Occasions per week of doing hobbies"
attributes(ah_wave1_v1_subset$TV)$label <- "Occasions per week of watching TV or playing videogames"
attributes(ah_wave1_v1_subset$Sports)$label <- "Occasions per week of playing sports"
attributes(ah_wave1_v1_subset$Exercise)$label <- "Occasions per week of doing exercise"

#send to RDS format
tmpdir <- Sys.getenv("TEMP")
saveRDS(object = ah_wave1_v1_subset, file = file.path(tmpdir, "ah_wave1_v1_subset.rds"))
ah_wave1_v1_subset_rds <- readRDS(file = file.path(tmpdir, "ah_wave1_v1_subset.rds"))
is(ah_wave1_v1_subset_rds)
ah_wave1_v1_subset_rds %>% 
    attributes() %>% 
    map(~ head(.))
```

We continue by showing some descriptive statistics on the weekly activities of Add Health study participants.  We chose to create a frequency table for each variable on time spent doing chores, doing hobbies, watching TV or playing video games, playing sports and getting exercise.
```{r, include =F}
#install.packages("captioner")
library(captioner)
table_nums <- captioner(prefix = "Table")
figure_nums <- captioner(prefix = "Figure")
# for having DIV tags
knitr::knit_hooks$set(class = function(before, options, envir) {
  if(before){
    sprintf("<div class = '%s'>", options$class)
  }else{
    "</div>"
  }
})

knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```
See `r table_nums(name = "tcap0", display = "cite")`

_`r table_nums(name = "tcap0", caption = "Time spent doing chores")`_
```{r}
#Frequency tables
#chores
chore_tab <-
  ah_wave1_v1_subset %>% 
    group_by(Chores) %>% 
    summarise(n = n()) %>% 
    mutate(`%` = n / sum(n) * 100) %>% 
    mutate(`%` = `%` %>% round(1)) %>% 
    kable(caption = "Participants' No. Occasions per week doing chores") %>% 
    kable_styling(full_width = FALSE, position = "left")
chore_tab
```
See `r table_nums(name = "tcap1", display = "cite")`

_`r table_nums(name = "tcap1", caption = "Time spent doing hobbies")`_
```{r}
#hobbies
hobby_tab <-
  ah_wave1_v1_subset %>% 
    group_by(Hobbies) %>% 
    summarise(n = n()) %>% 
    mutate(`%` = n / sum(n) * 100) %>% 
    mutate(`%` = `%` %>% round(1)) %>% 
    kable(caption = "Participants' No. Occasions per week doing hobbies") %>% 
    kable_styling(full_width = FALSE, position = "left")
hobby_tab
```
See `r table_nums(name = "tcap2", display = "cite")`

_`r table_nums(name = "tcap2", caption = "Time spent watching TV and playing video games")`_
```{r}
#TV and video games
tv_tab <-
  ah_wave1_v1_subset %>% 
    group_by(TV) %>% 
    summarise(n = n()) %>% 
    mutate(`%` = n / sum(n) * 100) %>% 
    mutate(`%` = `%` %>% round(1)) %>% 
    kable(caption = "Participants' No. Occasions per week watching TV or playing videogames") %>% 
    kable_styling(full_width = FALSE, position = "left")
tv_tab
```
See `r table_nums(name = "tcap3", display = "cite")`

_`r table_nums(name = "tcap3", caption = "Time spent playing sports")`_
```{r}
#sports
sport_tab <-
  ah_wave1_v1_subset %>% 
    group_by(Sports) %>% 
    summarise(n = n()) %>% 
    mutate(`%` = n / sum(n) * 100) %>% 
    mutate(`%` = `%` %>% round(1)) %>% 
    kable(caption = "Participants' No. Occasions per week playing sports") %>% 
    kable_styling(full_width = FALSE, position = "left")
sport_tab
```
See `r table_nums(name = "tcap4", display = "cite")`

_`r table_nums(name = "tcap4", caption = "Time spent getting exercise")`_
```{r}
#exercise
exercise_tab <-
  ah_wave1_v1_subset %>% 
    group_by(Exercise) %>% 
    summarise(n = n()) %>% 
    mutate(`%` = n / sum(n) * 100) %>% 
    mutate(`%` = `%` %>% round(1)) %>% 
    kable(caption = " Participants' No. Occasions per week doing exercise") %>% 
    kable_styling(full_width = FALSE, position = "left")
exercise_tab
```

We also created a frequency table that shows the amount of times per week participants in difference TV/videogame categories spend getting exercise.  See `r table_nums(name = "tcap6", display = "cite")`

See `r table_nums(name = "tcap6", display = "cite")`

_`r table_nums(name = "tcap6", caption = "Time spent getting exercise among participant with different TV and video game time use habits")`_
```{r}
#Frequency table with two variables
#exercise and tv
exercise_tv_tab <-
ah_wave1_v1_subset %>% 
    group_by(TV,
             Exercise) %>% 
    summarise(n = n(), .groups = "drop_last") %>% 
    mutate(`%` = n / sum(n) * 100) %>% 
    mutate(`%` = `%` %>% round(1)) %>% 
    kable(caption = "Participants' No. Occasions per week playing getting exercise by TV watching status") %>% 
    kable_styling(full_width = FALSE, position = "left")
exercise_tv_tab
```
Finally, we created a graph that displays the exercise frequency of survey participants that spend different amounts of time watching tv and playing video games. See _`r figure_nums(name = "Graph 1", display = "cite")`_ .
```{r}
#Graph
exercise_tv_plot <-
  ggplot(data = ah_wave1_v1_subset, mapping = aes(x = Exercise)) +
    geom_histogram(stat = "count") + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    labs(title=" Figure 1: Exercise frequency of survey participants that spend different amounts 
         of time watching tv and playing video games", subtitle="Data from 6504 people", 
       y="No. Participants", x="Times per week getting exercise", caption="Distribution is subset no. times participants watch tv and play video games") +
    facet_grid(~TV)
exercise_tv_plot 
```
\    
_`r figure_nums(name = "Graph 1", caption = "Exercise frequency of survey participants that spend different amounts of time watching tv and playing video games")`_.